Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-19 text/paper/revision_01/paper.tex
final edits together

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings compared to
each other (if they are detectable by empirical means at all)?
Addressing this question empirically is relevant because it lies at
the heart of a current debate about the nature of quantity inferences.
The controversy is usually seen as one between two main competing
theoretical positions, which we call \emph{traditionalism} and
\emph{grammaticalism} and take to be general ideas behind large
classes of approach to quantity implicatures \citep[c.f.][for related
discussion with a focus on deriving processing
theories]{ChemlaSingh2014:Remarks-on-the-}. Unfortunately, it is
rather difficult to empirically test either ``core theory''
(terminology from \citeauthor{ChemlaSingh2014:Remarks-on-the-})
directly because additional ``auxiliary assumptions'' are needed to
derive falsifiable \emph{ex ante} predictions about to-be-expected
observations in a concrete behavioral experiment, even irrespective of
processing accounts. We therefore try to carefully set up plausible
instantiations of either core theory with particular auxiliary
assumptions that give testable predictions at the outset in
Section~\ref{sec:theories-predictions}, then evaluate the obtained
predictions against our data, and finally broaden our perspective to
reflect in Section~\ref{sec:critical-reflection} on the impact of our
data on the more general tenability of the core theories.

But even with crisp theoretical predictions at hand, matters are still
very complex, as reflected by lively discussions about which
conclusions are permitted by previous experimental research
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The
results of these studies were discussed controversially because of
several specific reasons. Firstly, and most importantly, previous
studies presented target sentences visually.  However, traditionalists
often acknowledge the availability of local readings for prosodically
marked utterances
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under
this view, evidence for the availability of local readings
\citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}
may be an effect of “silent prosody” \citep[see
e.g.][]{Bader98,Fodor98}. We therefore address this position as the
\emph{prosodic markedness hypothesis} explicitly. Secondly, as
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences,
might well be confounded with, in particular, “typicality effects”
(see Section~\ref{sec:previous-studies} for discussion).  Finally,
previous studies have only accumulated limited evidence pertaining to
the relative salience of candidate readings of sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out plausible
\emph{ex ante} predictions of the competing core
theories. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} describes our experimental design,
analyses and results. Finally, Section~\ref{sec:critical-reflection}
discusses our method and interpretation of the results in a broader
theoretical context.

@@

There are two main theoretical positions which appear to make
different predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \emph{traditionalism} and \emph{grammaticalism}
and treat each in turn. Both approaches are addressed here as ``core
theories'': two main ideas behind several approaches to quantity
implicatures. To determine concrete predictions from these, additional
``auxiliary assumptions'' are necessary. Here, we spell a set of such
auxiliary assumptions for each core theory. Our motivation for the
particular choices we make is partly guided by practical concerns, so
as to have a concrete set of predictions for each core theory to
formulate expectations and compare our data to. We therefore introduce
some prominent or otherwise relevant variants of either core theory,
the predictions of which are summarized in Table~\ref{tab:predictions}
on page~\pageref{tab:predictions}. How other potential instantiations
of either core theory fare in light of our data will be discussed in
Section~\ref{sec:critical-reflection}.

Traditionalists frequently concede that local readings are licensed if
prosodically marked. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

@@

Strong and weak traditionalism differ with respect to whether the
assumptions needed to derive non-epistemic implicatures are prominent
in the given context of utterance (or some experiment or other). These
assumptions include relevance of the information in question, mutual
awareness that the stronger alternative has been a speaker option, the
speaker's competence about the issue, etc. \emph{Strong
  traditionalism} assumes that the additional assumptions needed to
derive non-epistemic implicatures are sufficiently salient to prefer
an implicature reading over a literal reading; \emph{weak
  traditionalism} assumes the reverse preference order. In other
words, in the context of a given experiment, it is compatible with a
traditionalist position to expect literal or implicature readings to
be more prominent. Or, put yet differently, a traditionalist could be
uncertain as to whether a given experiment is more likely to trigger
more implicature readings or literal readings. Without more
information about the experiment or context of interpretation,
traditionalism is compatible with either salience ordering.

@@

Luckily, the last kind of disambiguation approach does yield a clear
and non-trivial restriction of reading preferences off-the-shelf. It
selects readings by comparison of logical strengths, as suggested by
the \emph{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Supplementing
grammaticalism with a strength-based selection criterion has been
suggested, in one form or another, in many contributions to this
paradigm
\citep[e.g.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. At
the same time, the strongest meaning hypothesis has been frequently
called upon for disambiguation purposes in other domains as well,
including reciprocals \citep{DalrympleKanazawa1998:Reciprocal-Expr},
plural predication \citep{Winter2001:Plural-Predicat} or vagueness
\citep{CobrerosEgre2012:Tolerant-Classi}. For these reasons, we focus
from now on this latter type of selection
principle. Section~\ref{sec:critical-reflection} reconsiders
alternative selection principles.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation that make non-trivial predications
about preferred readings of \as- and \es-sentences (unlike, say, the
more detailed variants of \cite{FoxSpector:Economy-and-Emb}): fix a
sentence with propositional content $S$ and candidate readings $C(S)$,
where $C(S)$ contains $S$ and all readings derivable from inserting
$\exh(\cdot)$ at suitable scope sites;\fn{Strictly speaking, we should
  compare \emph{parses} of sentences with $\exh(\cdot)$ at various
  scope sites with respect to the readings that these parses give rise
  to \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may
  ignore this detail here for ease of exposition.} then define for
$X,Y \in C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\emph{weak grammaticalism} and \emph{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

\noindent The \emph{prosodic markedness hypothesis}, as we will call
it here, is then an addition to the traditionalist's explanatory
repertoire. Under this extra assumption, local readings are a marked
phenomenon, and not the product of the same process as run-of-the-mill
scalar implicatures. If supplemented with the prosodic markedness
hypothesis, traditionalists would therefore predict that local
readings are available for \as- and \es-sentences, but only if
prosodicly marked by contrastive stress.

To date, empirical investigations of this assumption are still scarce.  One
notable exception is a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternative
paraphrases  (such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  paper does not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect the
silent reading data, but raises the possibility that prosodic effects
may have been present at least in a subset of the items (see
footnote~1 in \citet{Frazier08}, p.~330).

@@

To sum up so far, though an effect of contrastive accentuation has
been claimed to be essential for triggering local implicatures,
experimental evidence is currently mixed, suggesting potential effects
of presentation mode (visual vs.~auditory), item-specific
characteristics, as well as the specific construction under
investigation (e.g., embedded cases vs.~non-embedded cases such as in
\citet{SchwarzClifton2008:Strengthening-o}). It is for these reasons
that an experimental investigation into the availability of readings
of \as- and \es-sentences should present sentential material
auditorily and explicitly address the prosodic markedness
hypothesis.

@@

We therefore included sentences with \emph{attachment ambiguities} as
in (\ref{bsp:target-related}), which have been shown to exhibit
interpretive preferences and can be disambiguated by prosodic
information.\footnote{It is possible that attachment ambiguities are
  different from the ``pragmatic ambiguity'' in \as- and
  \es-sentences, so that information about the former is not ideally
  informative about the latter. We will address this worry in
  Section~\textcolor{red}{XYZ}, where we argue that this does not
  affect the main conclusion that we would like to draw from our data
  eventually.}

@@

Judgements obtained for the four target conditions are listed in
Table~\ref{tab:count_data_CC}. We coded these answers as {\it
  literal}, {\it global} or {\it local} if they were as expected under
one of these readings and as {\it error} if not. The majority of
answers are indicative of a literal reading (\as-neutral: 90.4\%,
\as-accented: 86.8\%, \es-neutral: 68.4\%, \es-accented: 68.4\%). Only
two answers across all critical conditions fall into the category
indicative of global readings. Judgements indicative of local
readings, on the other hand, are more frequent (\as-neutral: 2.6\%,
\as-accented: 7.0\%, \es-neutral: 12.3\%, \es-accented: 14.0\%).

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are 7.0\%
/ 4.4\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.3\% / 17.6\% in the neutral / accented
\es-conditions respectively. A look at the left-hand side of
Table~\ref{tab:count_data_CC} reveals that there is some systematicity
in 'error' answers. Most answers that we classified as 'errors' have a
plausible explanation: random errors, a handful of mistakes resulting
from confusing buttons for \emph{true} and \emph{false} and a few
'spill-over errors' where participants gave responses one position too
late. There is one notable exception, though. A substantial number of
\emph{true} responses were given at the second position of
\es-sequences (9.7\% / 8.8\%) which is shown in
Figure~\ref{fig:exseqES2}. Although this position in the sequence is
the critical position to reveal a global reading, the latter would be
indicated by a \emph{false} response. It appears that subjects read
\es-sentences as if the non-monotonic quantifier \emph{exactly one}
was a monotonic \emph{at least one}, without pragmatic enrichment of
\emph{some}. 16 out of the total 21 answers in this category come from
5 subjects, while the remaining 5 were single answers by 5 different
subjects.
% subjects 1, 9, 12, 21, 84
This suggests that at least for some subjects non-monotonic
quantifiers were problematic. (We will come back to this intriguing
observation later in Section~\ref{sec:discussion-1}, where we suggest
that a non-monotonic reading of \emph{exactly one} might also explain
the high number of \emph{true} answers in step 5 of the
\es-conditions.)

@@

There are only very few answers that appear to be random
mistakes. Some answers in the `error' category are plausibly
`spill-over errors', where participants gave answers one step too
late. But on top of that, a substantial number of `error' answers
occurs on critical positions. For example, in the \lc-sequence there
are 23 \emph{true} answers at step 6 where \emph{false} answers would
be taken as indicative of \ec-readings. In the \ec-sequence there are
28 \emph{false} answers at step 5 where \emph{true} answers would be
taken as indicative of \lc-readings. Finally, prosodic cues for
dispreferred \ec-readings seem to have led to more mistakes than
other prosodic patterns.

@@

High scores of success in control conditions show that participants
understood the incremental verification task and were able to give
judgements at the right point in a picture sequence. Results for the
preference-related controls show that our design is able to detect
multiple readings and preferences among these. Crucially, the amount
of response types for \lc- and \ec-readings reflected the known
preference for the former, with a mild bias for readings that can be
judged later in the sequence. Also, the incremental verification task
was generally sensitive to prosody. However, significant prosodic
effects were restricted to preference-related controls. Contrastively
accenting scalar \emph{some} generally led to more local readings, but
that effect was not significant. Local readings occurred in
\es-sentences, independently of prosody, but required prosodic marking
in \as-sentences. Our analyses support the impression that inspection
of the raw data gives, namely that the general qualitative pattern of
reading preferences is the same for \as- and \es-conditions: literal
readings are preferred over local readings which are preferred over
global readings. More concretely, our analyses suggest the following
preference patterns (bracketed readings are not supported by our data
and the previous analyses):

@@

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemically weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The small number of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as for unaccented \es-sentences. It would also remain unclear why
prosodic cues should have such differential effects on \as- and
\es-sentences.

@@

\noindent Notice that $Alt_1$ is empty for \es-sentences if we
restrict attention to the ``global alternatives'' that asymmetrically
entail the to-be-interpreted sentence. Interestingly, the readings
that this approach derives are exactly the literal readings (under
empty alternatives) and the local readings (under non-empty
alternatives). What is left to explain is the preference relation over
these readings. Here, the Magri-style approach could resort to
considerations of \emph{parsing economy}: reasoning with fever
alternatives is easier than reasoning with more alternatives. This
would explain the preference for literal readings and also why local
readings are more strongly attested for \es-sentences.

@@

\paragraph{Traditionalism.} Traditionalism, as we described it in
Section~\ref{sec:theories-predictions}, is also uncomfortable with our
data. It has to explain the possibility of local readings of
prosodically unmarked \es-sentences, and, perhaps, also the large
number of local readings in accented \es-sentences compared to the
much lower number in accented \as-sentences. To do so within a strict
traditionalist framework seems difficult. A first option might be to
account for alleged local readings as the result of some other
phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

@@

The foregoing interpretation and discussion presupposed that the
interpretation of our data, given in Section~\ref{sec:discussion-1},
was sound. But there are a number of conceivable objections to our
design and our interpretation of the presented results, which
essentially all cast into doubt that our data suggests the preference
relation in (\ref{bsp:observed-preferences}). We will consider the
most pressing ones in the following.

@@

This objection is very serious. It is amplified by an even more
general worry, raised as a challenge by a reviewer, who intuits that
the incremental verification might altogether be insensitive to scalar
implicatures. According to the reviewer, it is to be expected that
since \emph{some} is typically strengthened in non-embedded
conditions, virtually all of our participants would exit at an early
position corresponding to the literal reading in a suitable version of
our \ivt.

>>>>

Given this plurality of hypothesized readings, the question arises:
how \emph{salient} are these theoretically conceivable readings compared to
each other, or, put differently, what is the \emph{preference order} on
hypothesized readings (if they are detectable by empirical means at all)?
Addressing this question empirically is relevant because it lies at
the heart of a current debate about the nature of quantity inferences.
The controversy is usually seen as one between two main competing
theoretical positions, which we call \emph{traditionalism} and
\emph{grammaticalism} and take to be general ideas behind large
classes of approaches to quantity implicatures \citep[c.f.][for related
discussion with a focus on deriving processing
theories]{ChemlaSingh2014:Remarks-on-the-}. % Unfortunately, it is
% rather difficult to empirically test either ``core theory''
% (terminology from \citeauthor{ChemlaSingh2014:Remarks-on-the-})
% directly because additional ``auxiliary assumptions'' are needed to
% derive falsifiable \emph{ex ante} predictions about to-be-expected
% observations in a concrete behavioral experiment, even irrespective of
% processing accounts. We therefore try to carefully set up plausible
% instantiations of either core theory with particular auxiliary
% assumptions that give testable predictions at the outset in
% Section~\ref{sec:theories-predictions}, then evaluate the obtained
% predictions against our data, and finally broaden our perspective to
% reflect in Section~\ref{sec:critical-reflection} on the impact of our
% data on the more general tenability of the core theories.

Results from previous empirical studies have been discussed
controversially
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. This
is because of several reasons. Firstly, and most importantly, previous
studies presented target sentences visually.  However, traditionalists
often acknowledge the availability of local readings for prosodically
marked utterances
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under
this view, evidence for the availability of local readings
\citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}
may be an effect of “silent prosody” \citep[see
e.g.][]{Bader98,Fodor98}. We therefore address this position as the
\emph{prosodic markedness hypothesis} explicitly. Secondly, as
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences,
might well be confounded with, in particular, “typicality effects”
(see Section~\ref{sec:previous-studies} for discussion).  Finally,
previous studies have only accumulated limited evidence pertaining to
the relative salience of candidate readings of sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out plausible
\emph{ex ante} predictions of \emph{traditionalism} and
\emph{grammaticalism}. Section~\ref{sec:previous-studies} reviews
related experimental studies to the extent necessary to motivate our
own approach. Section~\ref{sec:exp} describes our experimental design,
analyses and results. Finally, Section~\ref{sec:critical-reflection}
discusses our method and interpretation of the results in a broader
theoretical context.

@@

There are two main theoretical positions which appear to make
different predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \emph{traditionalism} and \emph{grammaticalism}
and treat each in turn. Both approaches are addressed here as ``core
theories'' (terminology from
\citeauthor{ChemlaSingh2014:Remarks-on-the-}): two main ideas behind
several approaches to quantity implicatures. To determine concrete
predictions from these, additional ``auxiliary assumptions'' are
necessary to derive \emph{ex ante} falsifiable predictions about
to-be-expected observations in a concrete behavioral experiment, even
irrespective of processing accounts. Here, we spell out a set of such
auxiliary assumptions for each core theory. Our motivation for the
particular choices we make is partly guided by practical concerns, so
as to have a concrete set of predictions for each core theory to
formulate expectations and compare our data to. We therefore introduce
certain relevant variants of either core theory, the predictions of
which are summarized in Table~\ref{tab:predictions} on
page~\pageref{tab:predictions}. How other potential instantiations of
either core theory fare in light of our data will be discussed in
Section~\ref{sec:critical-reflection}.

Traditionalists frequently concede that local readings are licensed if
prosodically marked. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief recap of the relevant experimental data.

@@

Strong and weak traditionalism differ with respect to whether the
assumptions needed to derive (non-epistemic) implicatures are
plausibly warranted in the given context of utterance (or some
experiment or other). These assumptions include relevance of the
information in question, mutual awareness that the stronger
alternative has been a speaker option, the speaker's competence about
the issue, etc. \emph{Strong traditionalism} assumes that the
additional assumptions needed to derive (non-epistemic) implicatures
are sufficiently salient to prefer an implicature reading over a
literal reading; \emph{weak traditionalism} assumes the reverse
preference order. In other words, in the context of a given
experiment, it is compatible with a traditionalist position to expect
literal or implicature readings to be more prominent. Or, put yet
differently, a traditionalist could be uncertain as to whether a given
experiment is more likely to trigger more implicature readings or
literal readings. Without more information about the experiment or
context of interpretation, traditionalism is compatible with either
salience ordering.

@@

The last kind of disambiguation approach does yield a clear
restriction of reading preferences for the case at hand. It selects
readings by comparison of logical strengths, as suggested by the
\emph{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Supplementing
grammaticalism with a strength-based selection criterion has been
suggested, in one form or another, in many contributions to this
paradigm
\citep[e.g.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. At
the same time, the strongest meaning hypothesis has been frequently
called upon for disambiguation purposes in other domains as well,
including reciprocals \citep{DalrympleKanazawa1998:Reciprocal-Expr},
plural predication \citep{Winter2001:Plural-Predicat} or vagueness
\citep{CobrerosEgre2012:Tolerant-Classi}. For these reasons, we focus
from now on this latter type of selection
principle. Section~\ref{sec:critical-reflection} reconsiders
alternative selection principles.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation that make non-trivial predictions
about preferred readings of \as- and \es-sentences (unlike, say, the
more detailed variants of \cite{FoxSpector:Economy-and-Emb}): fix a
sentence with propositional content $S$ and candidate readings $C(S)$,
where $C(S)$ contains $S$ and all readings derivable from inserting
$\exh(\cdot)$ at suitable scope sites;\fn{Strictly speaking, we should
  compare \emph{parses} of sentences with $\exh(\cdot)$ at various
  scope sites with respect to the readings that these parses give rise
  to \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may
  ignore this detail here for ease of exposition.} then define for
$X,Y \in C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\emph{weak grammaticalism} and \emph{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

\noindent The \emph{prosodic markedness hypothesis}, as we will call
it here, is then an addition to the traditionalist's explanatory
repertoire. Under this extra assumption, local readings are a marked
phenomenon, and not the product of the same process as run-of-the-mill
scalar implicatures. If supplemented with the prosodic markedness
hypothesis, traditionalists would therefore predict that local
readings are available for \as- and \es-sentences, but only if
prosodically marked by contrastive stress.

To date, empirical investigations of this assumption are still scarce.  One
notable exception is a pilot study reported in
\citet{Frazier08}, in which participants were instructed to read
embedded and non-embedded versions of English \as-sentences. In the
embedded sentences (e.g. {\it All of the students wrote some of the
  official memos}), {\it some} was either capitalized or
non-capitalized. Prior to the experiment, participants were informed
that capitalizing corresponded to contrastive stress. The task
consisted in a forced-choice questionnaire with two given alternative
paraphrases  (such as {\it All of the students wrote some but not all of the
  official memos} vs. {\it All of the students wrote at least some of
  the official memos}). Interestingly, strengthened interpretations
occurred rather often across all sentences in this task (59 \%), but
were not affected by capitalizing.\fn{Note that, unfortunately, the
  paper does not provide percentages for the embedded conditions
  alone, but only the means across sentence types.}
\citeauthor{Frazier08} concludes that accentuation did not affect the
silent reading data, but raises the possibility that prosodic effects
may have been present at least in a subset of the items (see
Footnote~1 in \citet{Frazier08}, p.~330).

@@

To sum up so far, though an effect of contrastive accentuation has
been claimed to be essential for traditionalism for triggering local
implicatures, experimental evidence is currently mixed, suggesting
potential effects of presentation mode (visual vs.~auditory),
item-specific characteristics, as well as the specific construction
under investigation (e.g., embedded cases vs.~non-embedded cases such
as in \citet{SchwarzClifton2008:Strengthening-o}). It is for these
reasons that an experimental investigation into the availability of
readings of \as- and \es-sentences should present sentential material
auditorily and explicitly address the prosodic markedness hypothesis.

@@

We therefore included sentences with \emph{attachment ambiguities} as
in (\ref{bsp:target-related}), which have been shown to exhibit
interpretive preferences and can be disambiguated by prosodic
information.\footnote{It is possible that attachment ambiguities are
  different from the ``pragmatic ambiguity'' in \as- and
  \es-sentences, so that information about the former is not ideally
  informative about the latter. We will address this worry in
  Section~\ref{sec:critical-reflection}, where we argue that this does
  not affect the main conclusion that we would like to draw from our
  data eventually.}

@@

Judgements obtained for the four target conditions are listed in
Table~\ref{tab:count_data_CC}. We coded these answers as {\it
  literal}, {\it global} or {\it local} if they were as expected under
one of these readings and as {\it error} if not. The majority of
answers are indicative of a literal reading (\as-neutral: 90.4\%,
\as-accented: 86.8\%, \es-neutral: 68.4\%, \es-accented: 68.4\%). Only
two answers (0.4\%) across all critical conditions fall into the category
indicative of global readings. Judgements indicative of local
readings, on the other hand, are more frequent (\as-neutral: 2.6\%,
\as-accented: 7.0\%, \es-neutral: 12.3\%, \es-accented: 14.0\%).

The number of responses that fell out of our classification scheme
differs substantially between \as- and \es-conditions. There are 7.0\%
/ 4.4\% unclassifiable judgements in the neutral / accented
\as-conditions, but 19.3\% / 17.6\% in the neutral / accented
\es-conditions respectively. A look at the left-hand side of
Table~\ref{tab:count_data_CC} reveals that there is some systematicity
in 'error' answers. Most answers that we classified as 'errors' have a
plausible explanation: random errors, a handful of mistakes resulting
from confusing buttons for \emph{true} and \emph{false} and a few
'spill-over errors' where participants gave responses one position too
late. There is one notable exception, though. A substantial number of
\emph{true} responses were given at the second position of
\es-sequences (9.7\% / 8.8\%) which is shown in
Figure~\ref{fig:exseqES2}. Although this position in the sequence is
the critical position to reveal a global reading, the latter would be
indicated by a \emph{false} response. Given the low number of possible
mistakes resulting from confusion buttons for \emph{true} and
\emph{false} elsewhere, this seems to be a more systematic pattern. It
appears as if subjects read \es-sentences as if the non-monotonic
quantifier \emph{exactly one} was a monotonic \emph{at least one},
without pragmatic enrichment of \emph{some}. 16 out of the total 21
answers in this category come from 5 subjects, while the remaining 5
were single answers by 5 different subjects.
% subjects 1, 9, 12, 21, 84
This suggests that at least for some subjects non-monotonic
quantifiers were problematic. (We will come back to this intriguing
observation later in Section~\ref{sec:discussion-1}, where we suggest
that a non-monotonic reading of \emph{exactly one} might also explain
the high number of \emph{true} answers in step 5 of the
\es-conditions.)

@@

There are only very few answers that appear to be random
mistakes. Some answers in the `error' category are plausibly
`spill-over errors', where participants gave answers one step too
late. But on top of that, a substantial number of `error' answers
occurs on the critical positions. For example, in the \lc-sequence
there are 23 \emph{true} answers at step 6 where \emph{false} answers
would be taken as indicative of \ec-readings. In the \ec-sequence
there are 28 \emph{false} answers at step 5 where \emph{true} answers
would be taken as indicative of \lc-readings. Finally, prosodic cues
for dispreferred \ec-readings seem to have led to more mistakes than
other prosodic patterns.

@@

High scores of success in control conditions show that participants
understood the incremental verification task and were able to give
judgements at the right point in a picture sequence. Results for the
preference-related controls show that our design is able to detect
multiple readings and preferences among these. Crucially, the amount
of response types for \lc- and \ec-readings reflected the known
preference for the former, with a mild bias for readings that can be
judged later in the sequence. Also, the incremental verification task
was generally sensitive to prosody. However, our regression analysis
showed that significant prosodic effects were restricted to
preference-related controls. Contrastively, the Bayesian analysis
showed that accenting scalar \emph{some} generally led to more local
readings, but that effect was not significant. Local readings occurred
in \es-sentences, independently of prosody, but required prosodic
marking in \as-sentences. Our analyses support the impression that
inspection of the raw data gives, namely that the general qualitative
pattern of reading preferences is the same for \as- and
\es-conditions: literal readings are preferred over local readings
which are preferred over global readings. More concretely, our
analyses suggest the following preference patterns (bracketed readings
are not supported by our data and the previous analyses):

@@

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemically weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The finding of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as for unaccented \es-sentences. It would also remain unclear why
prosodic cues should have such differential effects on \as- and
\es-sentences.

@@

\noindent Notice that $Alt_1$ is empty for \es-sentences if we
restrict attention to the ``global alternatives'' that asymmetrically
entail the to-be-interpreted sentence. Interestingly, the readings
that this approach derives are exactly the literal readings (under
empty alternatives) and the local readings (under non-empty
alternatives). What is left to explain is the preference relation over
these readings. Here, the Magri-style approach could resort to
considerations of \emph{economy}: reasoning with fever alternatives is
easier than reasoning with more alternatives. This would explain the
preference for literal readings and also why local readings are more
strongly attested for \es-sentences.

@@

\paragraph{Traditionalism.} Traditionalism, as we described it in
Section~\ref{sec:theories-predictions}, can also not account for the
observed pattern of results. It has to explain the possibility of
local readings of prosodically unmarked \es-sentences, and, perhaps,
also the large number of local readings in accented \es-sentences
compared to the much lower number in accented \as-sentences. To do so
within a strict traditionalist framework seems difficult. A first
option might be to account for alleged local readings as the result of
some other phenomenon, such as typicality or contrast
\citep{Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar,Tielvan-Tiel2014:Quantity-Matter}.
Since this ties into a critical reflection on our design, we will
enlarge on the possibility of typicality- and contrast-based effects
in our setting. Eventually, we argue that contrast-effects alone are
unlikely to explain the observed high number of local answers in
\es-sentences, but that a ``modern traditionalist'' explanation of the
answer pattern might be available if we allow for the possibility that
\emph{exactly one} gets an unexpected ``referential interpretation.''

@@

The foregoing interpretation and discussion presupposed that the
interpretation of our data, given in Section~\ref{sec:discussion-1},
was sound. But there are a number of conceivable objections to our
design and our interpretation of the presented results, which
essentially cast into doubt that our data suggests the preference
relation in (\ref{bsp:observed-preferences}). We will consider the
most pressing ones in the following.

@@

This objection is very serious. It is amplified by an even more
general worry, raised as a challenge by a reviewer, who intuits that
the incremental verification might altogether be insensitive to scalar
implicatures. According to such a view, it could be that since
\emph{some} is typically strengthened in non-embedded conditions,
virtually all of our participants would exit at an early position
corresponding to the literal reading in a suitable version of our
\ivt.
